\documentclass[a4paper, 12pt]{article}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{amsmath}

\newcommand{\be}{\begin{equation*}}
\newcommand{\ee}{\end{equation*}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\title{Industrial Organization, Fall 2016\\
Problem set 0: Matlab basics}
\date{}
\begin{document}
\maketitle

This homework is not graded. Its sole purpose is to give you some basic hands-on experience with Matlab. Feel free to skip this problem set if you are confident in your Matlab skills.


\section*{Practice writing own code}

This exercise uses a simulated panel with 10 years of annual data on 600 firms (\texttt{hw1data.csv}). Each observation contains the following variables:
\begin{description}
	\item[year] Year, $t$
	\item[id\_firm] Firm ID, $j$
	\item[logl] Logarithm of labor, $\ell$
	\item[logk] Logarithm of the capital stock, $k$
	\item[logq] Logarithm of output (in terms of value added, i.e. \textbf{there is no need to include intermediate materials in the production function}), $q$
\end{description}

The objective is to estimate the Cobb-Douglas production function
	\be
		q_{jt} = \beta_0 + \beta_kk_{jt} + \beta_\ell\ell_{jt} + e_{jt}
	\ee
	using pooled OLS (point estimates only; no standard errors).
\begin{itemize}	
	\item{First, implement OLS by minimizing the sum of squared residuals
	\be
		\min_{\beta_0, \beta_k, \beta_\ell}\sum_{j,t}\left(q_{jt} - \beta_0 - \beta_kk_{jt} - \beta_\ell\ell_{jt} \right)^2
	\ee
	That is, write a function that takes $\beta$'s, the vectors of $q_{jt}$, $k_{jt}$, $\ell_{jt}$ as input arguments and minimize it numerically w.r.t. $\beta$'s. In the language of tutorial 4, the data $q_{jt}$, $k_{jt}$ and $\ell_{jt}$ are the ``additional parameters''; check the above tutorial to find out how to pass data on to the objective function and minimize only w.r.t. $\beta$'s.
	
	You can use either \texttt{fminunc} or \texttt{fminsearch}. A good practice is to run both of them sequentially, using the solution from the former as a starting point for the latter. If all is well, the second solver should not venture too far from the starting point.}
	\item{The second way of implementing OLS is more efficient. Let $q$ be a column vector stacking all $q_{jt}$'s and $X = [1, k, \ell]$ be a $6000\times{3}$ matrix, made of ones, $k_{jt}$ and $\ell_{jt}$. Let $\beta = [\beta_0, \beta_k, \beta_\ell]'$ be the vector of coefficients. In this notation,
	\be
		\widehat\beta_{OLS} = \left(X'X\right)^{-1}X'q
	\ee
	Use this formula to find the estimates in one line of code. Make sure that the results match those obtained using the previous implementation of OLS.}
	\item{Implement the OLS estimator as a Matlab function that takes $q$ and $X = [1, k, \ell]$ as arguments (where $1$ is a vector of ones) and returns the estimate of $\beta = [\beta_0, \beta_k, \beta_\ell]'$.}
\end{itemize}

\end{document}